### 2020-06-27 covidWarehouse/manuscript/manuscript.Rmd
improve life table calculations + add occupancy risk factor

---
title: "Incidence of confirmed COVID-19 cases in care home staff and residents: Cohort study using electronic records in a large UK care home provider (March-June 2020)"
# author: "Dutey-Magni, PF, Williams, H, Hayward, A, Shallcross, L"
bibliography: references.bib
csl: elsevier-vancouver.csl
output:
  html_document:
    df_print: paged
    toc: yes
    code_folding: hide
  word_document:
    toc: yes
    reference_docx: template.docx
link-citations: yes
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
options(OutDec = "·")
knit_output_pdf <- F
if(knit_output_pdf){
  options(knitr.table.format = "pdf")
} else {
  options(knitr.table.format = "pandoc")
}

source("../.Rprofile")
devtools::load_all("..")
load_data()
library(dplyr)
library(lubridate)
library(kableExtra)
library(captioner)
library(survival)
library(survminer)
library(ggplot2)
library(finalfit)

knit_output_pdf <- F
if(knit_output_pdf){
  options(knitr.table.format = "pdf")
} else {
  options(knitr.table.format = "pandoc")
}

date_min <- as.Date("2020-03-02")
date_max <- as.Date("2020-06-07") #as.Date("2020-06-15")
date_min_tallies = as.Date("2020-03-24")
date_max_tallies = as.Date("2020-06-07")

```

```{r init_data}

tline <- dplyr::bind_rows(timelines$timeline)
stopifnot(unique(tline$date_end)>=date_max)
tline <- filter(tline, between(date, date_min, date_max)) %>% 
  mutate(home_type = factor(home_type, c("Residential", "Nursing"))) %>% 
  mutate(
    user_type2 = case_when(
      user_type %in% c("General", "Elderly") ~ "General/elderly",
      user_type == "Dementia" ~ "Dementia",
      TRUE ~ NA_character_ 
      ),
    admission_type2 = case_when(
      admission_type %in% c("Continuing Care", "Independent Living") ~ "Continuing care/independent living", 
      TRUE ~ admission_type
    ))

home_inclusion <- unique(tline$home_code)

reference_homes <- reference_homes %>% 
  filter(home_code %in% home_inclusion) %>% 
  left_join(select(filter(beds, year=="2020"), -year)) %>% 
  mutate(staff_to_bed_ratio = staff_total/beds,
         carer_to_bed_ratio = (staff_care_assistants + staff_nurses)/beds,
         bed_tot_category = cut(.$beds, 
           breaks = seq(20,100,25), right = F, 
           paste0(seq(20, 80, 25), "–", seq(34, 94, 25), " beds"))
         )

homes_country <- select(reference_homes, home_code, postcode) %>% 
  left_join(transmute(reference_geography, postcode, 
                      country, region_england = region), by = "postcode") %>% 
  mutate(region_UK = if_else(country == "England", region_england, country)) %>% 
  select(-postcode)
reference_homes <- left_join(reference_homes, homes_country)
rm(homes_country)

incidents <- incidents %>% 
  filter(dplyr::between(incident_date, date_min, date_max)) %>% 
  filter(home_code %in% home_inclusion)

desc_before_exclusions <- list()
desc_before_exclusions$nrow_incidents <- nrow(incidents)
desc_before_exclusions$missing_id <- sum(grepl("^MISSING", incidents$resident_id))
desc_before_exclusions$missing_id_pct <- mean(grepl("^MISSING", incidents$resident_id))*100
desc_before_exclusions$link_fail <- sum(!incidents$resident_id[
  which(!grepl("^MISSING", incidents$resident_id))] %in% unique(residents$resident_id))
desc_before_exclusions$link_fail_pct <- tbrounding(
  sum(!incidents$resident_id[
    which(!grepl("^MISSING", incidents$resident_id))] %in% unique(residents$resident_id)) / 
    nrow(incidents)*100)

# removing erroneous incident reports (no identifier or not linking to residents file) 
incidents <- filter(incidents, !grepl("^MISSING", resident_id)) %>% 
  filter(resident_id %in% unique(residents$resident_id))

new_cases <- filter(new_cases, between(date, date_min, date_max_tallies)) %>% 
    filter(home_code %in% home_inclusion)

approx_occupancy <- tline %>% 
  group_by(home_code, date) %>% 
  summarise(
    occupancy = sum(tidyr::replace_na(rday, 0)),
    susceptible_symptomatic = sum(susceptible_symptomatic, na.rm = T),
    susceptible_confirmed = sum(susceptible_confirmed, na.rm = T))
approx_occupancy <- merge(
  tibble(date = seq(date_min, date_max, 1)),
  tibble(home_code = home_inclusion),
  all = TRUE) %>% 
  left_join(approx_occupancy)

occupancy_interpolated <- resident_days_linear_approx() %>% 
  # TODO: currently putting the individual-level occup for 360 because no occ tally is known!
  # it's got 57 bed including 5 contract beds, so 
  left_join(transmute(filter(approx_occupancy, home_code == "360"), 
                      date, home_code  , occup360 = trunc(occupancy * 57/52)),
            by = c("home_code", "date")) %>%
  mutate(occupancy = if_else(home_code == "360", occup360, occupancy)) %>% 
  select(-occup360)

home_mean_occupancy <- occupancy_interpolated  %>% 
  filter(between(date, date_min_tallies, date_max_tallies)) %>% 
  group_by(home_code) %>% 
  summarise(mean_occupancy = mean(occupancy, na.rm = T))

reference_homes <- reference_homes %>% 
  left_join(home_mean_occupancy) %>% 
  mutate(pct_occupancy = mean_occupancy/beds)

residents_episodes <- tline %>% 
  filter(rday==1) %>% 
  group_by(resident_id, home_code, dob, gender, 
           status, home_type, user_type2,
           susceptible_symptomatic, 
           susceptible_confirmed) %>% 
  summarise(entry_date = min(date),
            exit_date = max(date+1),
            resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed)) %>% 
  mutate(date1 = as.numeric( entry_date - date_min),
         date2 = as.numeric(exit_date - date_min),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  mutate(symptomatic = -susceptible_symptomatic+1,
         confirmed = -susceptible_confirmed+1,
         symptomatic_confirmed = (-susceptible_symptomatic+1)*(-susceptible_confirmed+1),
         date2 = if_else(date1 == date2, date2 + 1, date2)) %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Negative",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Negative", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed")),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95))   %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category, beds, pct_occupancy)) %>% 
  group_by(resident_id) %>% 
  mutate(death_status = if_else(
    date1 == max(date1),
    as.integer(status == "Deceased"),
    0L
  )) %>% 
  ungroup()

residents_overview <- tline %>% 
  filter(rday==1) %>%
  mutate(ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  group_by(resident_id, home_code, ageg, gender, 
           admission_type2, home_type, user_type2, status) %>% 
  summarise(resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed),
            symptomatic = max(-susceptible_symptomatic+1, na.rm = T),
            confirmed = max(-susceptible_confirmed+1, na.rm = T),
            start_date = min(date, na.rm = T),
            stop_date = max(date, na.rm = T),
            start_int = as.numeric(min(date, na.rm = T) - date_min),
            stop_int = as.numeric(max(date, na.rm = T)- date_min)
    ) %>% 
  ungroup() %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Negative",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Negative", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed"))) %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category, beds, beds, pct_occupancy)) %>% 
  left_join(distinct(select(incidents, resident_id, starts_with("covid_first"))))

# CARE HOME AGGREGATES
home_total_confirmed_DAT <- filter(residents_overview, !is.na(covid_first_confirmed)) %>% 
  filter(between(covid_first_confirmed, date_min_tallies, date_max_tallies)) %>% 
  group_by(home_code) %>% 
  summarise(DAT_confirmed = tidyr::replace_na(n_distinct(resident_id), 0))
home_total_confirmed_CT <- new_cases %>% group_by(home_code) %>% 
  summarise(CT_confirmed = sum(tidyr::replace_na(new_confirmed_home, 0) + 
                                 tidyr::replace_na(new_confirmed_hospital, 0)))

res_counts <- residents_overview %>% 
  group_by(home_code) %>% 
  summarise(unique_residents = n_distinct(resident_id))

desc_homes <- reference_homes %>% 
  left_join(res_counts, by = "home_code") %>%
  mutate(Scope = paste("FSHCG", country)) %>% 
  group_by(Scope) %>% 
  summarise(., 
            `Total homes` = n_distinct(home_code),
            `Total beds` = sum(beds),
            `Total contract beds` = sum(contract_beds, na.rm = T),
            `Total residents (excl. contract beds)` = sum(unique_residents, na.rm = T))

desc_national <- tibble(
  Scope = c("England", "Scotland (2017)", "Northern Ireland"),
  `Total homes` = c(9400, 1142, 483),
  `Total beds` = c(45000, 40926, 16095),
  `Total contract beds` = NA_integer_,
  `Total residents` = c(NA, 35989, NA)
)

home_total_confirmed <- distinct(reference_homes, home_code, home_name_clickview, postcode) %>% 
  left_join(select(reference_geography, postcode, country, region_UK), by = "postcode") %>% 
  left_join(res_counts, by = "home_code") %>% 
  left_join(home_total_confirmed_CT) %>% 
  left_join(home_total_confirmed_DAT) %>% 
  mutate(CT_confirmed = tidyr::replace_na(CT_confirmed, 0),
        DAT_confirmed = tidyr::replace_na(DAT_confirmed, 0))

home_zero_confirmed <- home_total_confirmed %>% 
  group_by(region_UK) %>% 
  summarise(`Total homes` = n(),
            zero_datix = sum(DAT_confirmed == 0),
            pct_datix = sum(DAT_confirmed == 0)/n(),
            zero_CT = sum(CT_confirmed == 0) ,
            pct_CT =  sum(CT_confirmed == 0)/n()) %>% 
  arrange(-pct_datix)

desc_homes <- bind_rows(desc_homes, desc_national)

```

```{r km_objects}

residents_ct <- tibble(date = seq(date_min, date_max_tallies, 1)) %>% 
  left_join(distinct(new_cases, date, home_code, 
                     new_confirmed_home, new_confirmed_hospital,
                     new_symptomatic_home, new_symptomatic_hospital)) %>% 
  filter(!is.na(home_code)) %>% 
  left_join(distinct(reference_homes, home_code, home_name_datix, postcode, staff_total, nursing_dementia, nursing_general, residential_dementia, residential_general, younger_person)) %>% 
  left_join(select(reference_geography, postcode, region_UK)) %>% 
  left_join(occupancy_interpolated) %>% 
  arrange(home_code, date) %>% 
  group_by(home_code) %>% 
  mutate(new_symptomatic = tidyr::replace_na(new_symptomatic_home, 0) +
                                    tidyr::replace_na(new_symptomatic_hospital, 0),
         new_confirmed = tidyr::replace_na(new_confirmed_home, 0) +
                                    tidyr::replace_na(new_confirmed_hospital, 0)) %>% 
  mutate(I_symptomatic = total_cases_in_home(occupancy, new_symptomatic, date),
         I_confirmed = total_cases_in_home(occupancy, new_confirmed, date)) %>% 
  ungroup()

km_ct_residents_confirmed <- residents_ct %>% 
  group_by(date, region_UK) %>% 
  summarise(
    confirmed = sum(new_confirmed),
    I_confirmed = round(sum(I_confirmed)),
    occupancy = sum(occupancy)
  ) %>% 
  group_by(region_UK) %>% 
  km_ct_estimator(df = ., t = date, d = confirmed, 
                  pop = (occupancy - I_confirmed), subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog)
  )

# residents_ct_symp <- tibble(date = seq(date_min, date_max_tallies, 1)) %>% 
#   left_join(distinct(new_cases, date, home_code, 
#                      new_symptomatic_home, new_symptomatic_hospital)) %>% 
#   filter(!is.na(home_code)) %>% 
#   left_join(distinct(reference_homes, home_code, home_name_datix, postcode)) %>% 
#   left_join(select(reference_geography, postcode, region_UK)) %>% 
#   left_join(occupancy_interpolated) %>% 
#   arrange(date)

km_ct_residents_symp <- residents_ct %>% 
  group_by(date, region_UK) %>% 
  summarise(
    symptomatic = sum(new_symptomatic),
    I_symptomatic = round(sum(I_symptomatic)),
    occupancy = sum(occupancy)
  ) %>% 
  group_by(region_UK) %>% 
  km_ct_estimator(df = ., t = date, d = symptomatic, 
                  pop = (occupancy - I_symptomatic), subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog)
  )

staff_ct <- tibble(date = seq(date_min, date_max_tallies, 1)) %>% 
  left_join(distinct(new_cases, date, home_code, 
                     new_colleague_confirmed, new_colleague_symptomatic)) %>% 
  filter(!is.na(home_code)) %>% 
  left_join(distinct(reference_homes, home_code, home_name_datix, postcode, staff_total, nursing_dementia, nursing_general, residential_dementia, residential_general, younger_person)) %>% 
  left_join(select(reference_geography, postcode, region_UK))  %>% 
  arrange(date) %>% 
  group_by(home_code) %>% 
  mutate(
    new_colleague_confirmed = tidyr::replace_na(new_colleague_confirmed, 0),
    new_colleague_symptomatic = tidyr::replace_na(new_colleague_symptomatic, 0)
    ) %>% 
  mutate(I_symptomatic = total_cases_in_home(staff_total, new_colleague_symptomatic, date),
         I_confirmed = total_cases_in_home(staff_total, new_colleague_confirmed, date)) %>% 
  ungroup()

km_staff_symptomatic <- staff_ct %>% 
  group_by(date, region_UK) %>% 
  summarise(
    symptomatic = sum(new_colleague_symptomatic),
    I_symptomatic = round(sum(I_symptomatic)),
    staff_total = sum(staff_total)
  ) %>% 
  group_by(region_UK) %>% 
  km_ct_estimator(., t=date, d=symptomatic, pop=(staff_total-I_symptomatic),
                  subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog)
  )

km_staff_confirmed <- staff_ct %>% 
  group_by(date, region_UK) %>% 
  summarise(
    confirmed = sum(new_colleague_confirmed),
    I_confirmed = round(sum(I_confirmed)),
    staff_total = sum(staff_total)
  ) %>% 
  group_by(region_UK) %>% 
  km_ct_estimator(., t=date, d=confirmed, pop=(staff_total-I_confirmed),
                  subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog)
  )

desc_staff <- list()
desc_staff$total_staff <- reference_homes %>% filter(home_code %in% unique(residents_overview$home_code)) %>% .$staff_total %>% sum()

desc_staff$CT_symptomatic_total <- sum(new_cases$new_colleague_symptomatic, na.rm = T)
desc_staff$CT_symptomatic_denom <- sum(km_staff_symptomatic$n_t)

desc_staff[paste0(
  "CT_symptomatic_p", c("", "_lower", "_upper")
)] <- epitools::binom.exact(
  desc_staff$CT_symptomatic_total,
  desc_staff$total_staff)[c("proportion", "lower", "upper")] *100

desc_staff[paste0(
  "CT_symptomatic_IR", c("", "_lower", "_upper")
)] <- epitools::pois.exact(
  desc_staff$CT_symptomatic_total,
  desc_staff$CT_symptomatic_denom)[c("rate", "lower", "upper")] *100000

desc_staff$CT_confirmed_total <- sum(new_cases$new_colleague_confirmed, na.rm = T)
desc_staff$CT_confirmed_denom <- sum(km_staff_confirmed$n_t)
desc_staff[paste0(
  "CT_confirmed_p", c("", "_lower", "_upper")
)] <- epitools::binom.exact(
  desc_staff$CT_confirmed_total,
  desc_staff$total_staff)[c("proportion", "lower", "upper")] *100

desc_staff[paste0(
  "CT_confirmed_IR", c("", "_lower", "_upper")
)] <- epitools::pois.exact(
  desc_staff$CT_confirmed_total,
  desc_staff$CT_confirmed_denom)[c("rate", "lower", "upper")] *100000

```

```{r descriptives}
desc_residents <- list()
desc_residents$DAT_total_residents <- dplyr::n_distinct(residents_overview$resident_id)
desc_residents$DAT_total_residents_england <- dplyr::filter(residents_overview, country == "England") %>% summarise(n = dplyr::n_distinct(resident_id)) %>% .$n

#COUNTS
# This is estimated in validation.Rmd
desc_residents$CT_total_residents <- round(desc_residents$DAT_total_residents * 1.117309 + .49)
desc_residents$CT_total_residents_england <- round(desc_residents$DAT_total_residents_england * 1.0793268 + .49)
desc_residents$CT_total_symptomatic <- sum(new_cases$new_symptomatic_home, na.rm=T) + 
  sum(new_cases$new_symptomatic_hospital, na.rm=T)
desc_residents$CT_total_symptomatic_denom <- sum(km_ct_residents_symp$n_t)
desc_residents[paste0(
  "CT_p_symptomatic", c("", "_lower", "_upper")
)] <- epitools::binom.exact(desc_residents$CT_total_symptomatic,
                            desc_residents$CT_total_residents)[
                              c("proportion", "lower", "upper")
                            ] * 100

desc_residents[paste0(
  "CT_IR_symptomatic", c("", "_lower", "_upper")
)] <- epitools::pois.exact(desc_residents$CT_total_symptomatic,
                            desc_residents$CT_total_symptomatic_denom)[
                              c("rate", "lower", "upper")
                            ] * 100000

desc_residents$CT_total_confirmed <-  sum(new_cases$new_confirmed_home, na.rm = T) + sum(new_cases$new_confirmed_hospital, na.rm = T)
desc_residents$CT_total_confirmed_denom <- sum(km_ct_residents_confirmed$n_t)

desc_residents[paste0(
  "CT_p_confirmed", c("", "_lower", "_upper")
)] <- epitools::binom.exact(desc_residents$CT_total_confirmed,
                            desc_residents$CT_total_residents)[
                              c("proportion", "lower", "upper")
                            ] * 100

desc_residents[paste0(
  "CT_IR_confirmed", c("", "_lower", "_upper")
)] <- epitools::pois.exact(desc_residents$CT_total_confirmed,
                            desc_residents$CT_total_confirmed_denom)[
                              c("rate", "lower", "upper")
                            ] * 100000

#INCIDENT REPORTS
desc_residents$unique_residents_incidents <- dplyr::n_distinct(incidents$resident_id)

desc_residents$DAT_total_symptomatic <- n_distinct(filter(incidents, !is.na(covid_first_symptomatic))$resident_id)
desc_residents$DAT_total_symptomatic_denom <- sum(residents_overview$susceptible_symptomatic_days)
desc_residents[paste0(
  "DAT_p_symptomatic", c("", "_lower", "_upper")
)] <- epitools::binom.exact(desc_residents$DAT_total_symptomatic,
                            desc_residents$DAT_total_residents)[
                              c("proportion", "lower", "upper")
                            ] * 100
desc_residents[paste0(
  "DAT_IR_symptomatic", c("", "_lower", "_upper")
)] <- epitools::pois.exact(desc_residents$DAT_total_symptomatic,
                            desc_residents$DAT_total_symptomatic_denom)[
                              c("rate", "lower", "upper")
                            ] * 100000

desc_residents$total_tests <- nrow(dplyr::filter(incidents, covid_tested == 1))
desc_residents$unique_residents_tested <- dplyr::filter(incidents, covid_tested == 1) %>% 
  dplyr::summarise(n = n_distinct(resident_id)) %>% .$n
desc_residents$unique_residents_test_result <- dplyr::filter(incidents, !is.na(covid_test_result)) %>% 
  dplyr::summarise(n = n_distinct(resident_id)) %>% .$n
desc_residents$unique_residents_test_result_positive <- dplyr::filter(incidents, covid_test_result == 1) %>%  
  dplyr::summarise(n = n_distinct(resident_id)) %>% .$n

desc_residents$DAT_total_confirmed_denom <- sum(residents_overview$susceptible_confirmed_days)
desc_residents[paste0(
  "DAT_p_confirmed", c("", "_lower", "_upper")
)] <- epitools::binom.exact(desc_residents$unique_residents_test_result_positive,
                            desc_residents$DAT_total_residents)[
                              c("proportion", "lower", "upper")
                            ] * 100

desc_residents[paste0(
  "DAT_IR_confirmed", c("", "_lower", "_upper")
)] <- epitools::pois.exact(desc_residents$unique_residents_test_result_positive,
                            desc_residents$DAT_total_confirmed_denom)[
                              c("rate", "lower", "upper")
                            ] * 100000

# CASE FATALITY AND RELATIVE RISK

desc_residents$DAT_total_deaths <- sum(residents_overview$status == "Deceased")
desc_residents$total_denom <- sum(residents_overview$resident_days)
desc_residents[paste0(
  "DAT_p_death",  c("", "_lower", "_upper")
)] <- epitools::binom.exact(
  desc_residents$DAT_total_deaths,
  desc_residents$DAT_total_residents
)[c("proportion", "lower", "upper")] *100

desc_residents[paste0(
  "DAT_IR_death",  c("", "_lower", "_upper")
)] <- epitools::pois.exact(
  desc_residents$DAT_total_deaths,
  desc_residents$total_denom
)[c("rate", "lower", "upper")] * 100000

desc_residents$DAT_CFR <- residents_overview %>% 
  filter(confirmed == 1) %>% 
  summarise(n = n(),
            y = sum(status == "Deceased"))

desc_residents$DAT_BFR <- residents_overview %>%
  filter(confirmed != 1) %>%
  summarise(n = n(),
            y = sum(status == "Deceased"))

desc_residents$DAT_CFR_RR <- relative_risk(desc_residents$DAT_CFR$y, desc_residents$DAT_CFR$n - desc_residents$DAT_CFR$y, desc_residents$DAT_BFR$y, desc_residents$DAT_BFR$n - desc_residents$DAT_BFR$y)

desc_residents$DAT_CFR_RR_SE <- relative_risk_SE(desc_residents$DAT_CFR$y, desc_residents$DAT_CFR$n - desc_residents$DAT_CFR$y, desc_residents$DAT_BFR$y, desc_residents$DAT_BFR$n - desc_residents$DAT_BFR$y)

desc_residents[paste0(
  "DAT_CFR", c("", "_lower", "_upper")
)] <- epitools::binom.exact(
  desc_residents$DAT_CFR$y,
  desc_residents$DAT_CFR$n
)[c("proportion", "lower", "upper")] *100

desc_residents[paste0(
  "DAT_BFR", c("", "_lower", "_upper")
)] <- epitools::binom.exact(
  desc_residents$DAT_BFR$y,
  desc_residents$DAT_BFR$n
)[c("proportion", "lower", "upper")] *100

### Homes
desc_residents$total_homes <- n_distinct(home_total_confirmed$home_code)
desc_residents$CT_zero_case <- sum(home_total_confirmed$CT_confirmed == 0)
desc_residents$DAT_zero_case <- sum(home_total_confirmed$DAT_confirmed == 0)

## OUTBREAK HOMES 
desc_residents_outbreak_only <- list()
desc_residents_outbreak_only$outbreak_homes <- home_total_confirmed$home_code[home_total_confirmed$CT_confirmed != 0]

desc_residents_outbreak_only$CT_total_beds <- beds %>% 
  filter(year == 2020 & home_code %in% desc_residents_outbreak_only$outbreak_homes) %>% 
  summarise(beds = sum(beds),
            contract_beds = sum(contract_beds))

desc_residents_outbreak_only$CT_total_residents <- residents_overview %>% 
  dplyr::filter(home_code %in% desc_residents_outbreak_only$outbreak_homes) %>% 
  dplyr::summarise(total_residents = round(dplyr::n_distinct(resident_id)* 1.1151324 + .49)) %>% 
  .$total_residents

desc_residents_outbreak_only$CT_total_confirmed <- new_cases %>% 
  filter(home_code %in% desc_residents_outbreak_only$outbreak_homes) %>% 
  summarise(n = sum(new_confirmed_home, na.rm = T) + sum(new_confirmed_hospital, na.rm = T)) %>% 
  .$n

desc_residents_outbreak_only$CT_total_symptomatic <- new_cases %>% 
  filter(home_code %in% desc_residents_outbreak_only$outbreak_homes) %>% 
  summarise(n = sum(new_symptomatic_home, na.rm = T) + sum(new_symptomatic_hospital, na.rm = T)) %>% 
  .$n

desc_residents_outbreak_only$CT_total_confirmed_denom <- residents_ct %>% 
  filter(home_code %in% desc_residents_outbreak_only$outbreak_homes) %>% 
  summarise(n = sum(occupancy - I_confirmed)) %>% 
  .$n

desc_residents_outbreak_only[paste0(
  "CT_p_confirmed", c("", "_lower", "_upper")
)] <- epitools::binom.exact(desc_residents_outbreak_only$CT_total_confirmed,
                            desc_residents_outbreak_only$CT_total_residents)[
                              c("proportion", "lower", "upper")
                              ] * 100

desc_residents_outbreak_only[paste0(
  "CT_IR_confirmed", c("", "_lower", "_upper")
)] <- epitools::pois.exact(desc_residents_outbreak_only$CT_total_confirmed,
                           desc_residents_outbreak_only$CT_total_confirmed_denom)[
                             c("rate", "lower", "upper")
                             ] * 100000

desc_residents_outbreak_only$CT_total_symptomatic <-  new_cases %>% 
  filter(home_code %in% desc_residents_outbreak_only$outbreak_homes) %>% 
  summarise(n = sum(new_symptomatic_home, na.rm = T) + sum(new_symptomatic_hospital, na.rm = T)) %>% 
  .$n

desc_residents_outbreak_only$CT_total_symptomatic_denom <- residents_ct %>% 
  filter(home_code %in% desc_residents_outbreak_only$outbreak_homes) %>% 
  summarise(n = sum(occupancy - I_symptomatic)) %>% 
  .$n

desc_residents_outbreak_only[paste0(
  "CT_p_symptomatic", c("", "_lower", "_upper")
)] <- epitools::binom.exact(desc_residents_outbreak_only$CT_total_symptomatic,
                            desc_residents_outbreak_only$CT_total_residents)[
                              c("proportion", "lower", "upper")
                            ] * 100
desc_residents_outbreak_only[paste0(
  "CT_IR_symptomatic", c("", "_lower", "_upper")
)] <- epitools::pois.exact(desc_residents_outbreak_only$CT_total_symptomatic,
                            desc_residents_outbreak_only$CT_total_symptomatic_denom)[
                              c("rate", "lower", "upper")
                            ] * 100000

```

```{r captions}
## TABLES
table_nums <- captioner::captioner(prefix = "Table")
tab_national_descriptives <- table_nums(
  name = "tab_national_descriptives",
  caption = "Total homes, beds, contract beds, and residents within FSHCG and nationwide")
tab_resident_descriptives <- table_nums(
  name = "tab_resident_descriptives",
  caption = paste0("Breakdown of FSHCG private bed residents by care home type, sex, age, region and status on study exit (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
supp_epi_CT <- table_nums(
  name = "supp_epi_CT",
  caption =  paste0("Incidence proportions and rates of SARS-CoV-2 infections in residents and staff according to FSHCG manager counts (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
supp_epi_CT_oubreak_only <- table_nums(
  name = "supp_epi_CT_oubreak_only",
  caption =  paste0("Incidence proportions and rates of SARS-CoV-2 infections among residents of outbreak care homes according to FSHCG manager counts (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
tab_CoxPH_confirmed <- table_nums(
  name = "tab_CoxPH_confirmed",
  caption =  paste0("Factors associated with the hazard of become a confirmed case: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))
tab_CoxPH_death <- table_nums(
  name = "tab_CoxPH_death",
  caption =  paste0("Factors associated with all-cause mortality: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))

supp_epi_DAT <- table_nums(
  name = "supp_epi_DAT",
  caption =  paste0("Incidence proportions and rates of SARS-CoV-2 infections amongst residents according to Datix incident reports (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))

tab_zero_cases <- table_nums(
  name = "tab_zero_cases",
  caption = paste0("Number and proportion of homes without confirmed cases by UK region (", format(date_min_tallies, "%d/%m/%y"), "-", format(date_max_tallies, "%d/%m/%y"), ")")
)

## FIGURES
fig_nums <- captioner(prefix = "Figure")

fig_KM_CT_residents_symptomatic <- fig_nums(
  name = "fig_KM_CT_residents_symptomatic",
  caption = paste0("Kaplan-Meier estimates of the cumulative probability of residents becoming symptomatic based on care home manager daily counts")
)
fig_KM_residents_confirmed <- fig_nums(
  name = 'fig_KM_residents_confirmed',
  caption = paste0("Kaplan-Meier estimates of the cumulative probability of residents (n=",
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ") becoming confirmed cases based on Datix incidents report"))
fig_KM_CT_confirmed_residents <- fig_nums(
  name = 'fig_KM_CT_confirmed_residents',
  caption = "Kaplan-Meier estimates of the cumulative probability of residents becoming confirmed cases based on care home manager daily counts")
fig_KM_mortality <- fig_nums(
  "fig_KM_mortality",
  caption = paste0("Kaplan-Meier estimates of resident (n=",
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ") survival by SARS-COV-2 case type"))
fig_KM_staff_symptomatic <- fig_nums(
  name = 'fig_KM_staff_symptomatic',
  caption = paste0("Kaplan-Meier estimates of the cumulative probability of care home staff (n=",   formatbm(desc_staff$total_staff),
  ") becoming symptomatic cases based on manager-reported daily counts"))
fig_KM_staff_confirmed <- fig_nums(
  name = 'fig_KM_staff_confirmed',
  caption = paste0("Kaplan-Meier estimates of the cumulative probability of care home staff (n=",   formatbm(desc_staff$total_staff),
  ") becoming confirmed cases based on manager-reported daily counts"))

```

```{r period_prevalence}
cis_data <- list(
  n_positive = 66,
  n_negative = 60618
)

england_results <- list()
england_results$n_positive = sum(inner_join(new_cases, filter(reference_homes, country == "England"))[c("new_confirmed_home", "new_confirmed_hospital")], na.rm = T)
england_results$n_negative = desc_residents$CT_total_residents_england  - england_results$n_positive

england_results$RR_england <- relative_risk(
  a = england_results$n_positive,
  b = england_results$n_negative,
  c = cis_data$n_positive,
  d = cis_data$n_negative)

england_results$RR_SE <- relative_risk_SE(
  a = england_results$n_positive,
  b = england_results$n_negative,
  c = cis_data$n_positive,
  d = cis_data$n_negative)
```

# Abstract

**Background** COVID-19 has caused substantial morbidity and mortality in care home residents and staff. National surveys suggest infection is declining in the general population in the UK, but outbreaks persist in care homes. Epidemiological data on disease trends in care home staff and residents are scarce. We used administrative data from the Four Seasons Health Care Group (FSHCG), a national care home chain representing 9% of all beds in England, Scotland and Northern Ireland.

**Methods:** Retrospective cohort study using individual-level electronic records from residents (incidents reports) and daily care home counts of infection for both residents and staff. We estimated the incidence proportion and incidence rate of symptoms, confirmed infection, and death in residents and staff, and compare them with findings from a national community survey. We tested a range of risk factors for resident infections in multivariate regression models.

**Findings:** An estimated `r formatbm(desc_residents$CT_total_residents)` residents were present in `r dplyr::n_distinct(residents$home_code)` FSHCG homes between `r format(date_min_tallies, "%e %B %Y")` and `r format(date_max_tallies, "%e %B %Y")`. During the period `r formatbm(desc_residents$CT_total_symptomatic_denom)` residents developed COVID-19 symptoms (`r tbrounding(desc_residents$CT_p_symptomatic)`% [95% confidence interval: `r tbrounding(desc_residents$CT_p_symptomatic_lower)`%; `r tbrounding(desc_residents$CT_p_symptomatic_upper)`%]), and `r formatbm(desc_residents$CT_total_confirmed)` (`r tbrounding(desc_residents$CT_p_confirmed)`% [`r tbrounding(desc_residents$CT_p_confirmed_lower)`%; `r tbrounding(desc_residents$CT_p_confirmed_upper)`%]) had an infection confirmed with nasopharyngeal RNA tests. Across all care homes, the confirmed infection rate of incidence was `r tbrounding(desc_residents$CT_IR_confirmed)` per 100,000 resident-days [`r tbrounding(desc_residents$CT_IR_confirmed_lower)`; `r tbrounding(desc_residents$CT_IR_confirmed_upper)`]. In England, `r england_results$n_positive` infections were confirmed of a total `r formatbm(england_results$n_positive + england_results$n_negative)` residents. Comparing with the community, the relative risk of confirmed cases in England was `r tbrounding(england_results$RR_england)` [`r tbrounding(exp(log(england_results$RR_england) - 1.96 * england_results$RR_SE)) `; `r tbrounding(exp(log(england_results$RR_england) + 1.96 * england_results$RR_SE))`]. These estimates include `r desc_residents$CT_zero_case`/`r desc_residents$total_homes` homes (`r tbrounding( desc_residents$CT_zero_case/desc_residents$total_homes*100)`%) that had not reported a single confirmed case.

The staff-to-bed ratio was the only factor to be independently associated with the rate of confirmed cases: an increase by 10 percentage point in the number of staff relative to the bed total was associated with a 13% reduction [9%, 18%] in the rate of confirmed cases. Out of `r formatbm(desc_staff$total_staff)` staff, `r formatbm(desc_staff$CT_confirmed_total)` confirmed infections were counted(
`r tbrounding(desc_staff$CT_confirmed_p)`% [`r tbrounding(desc_staff$CT_confirmed_p_lower)`%; `r tbrounding(desc_staff$CT_confirmed_p_upper)`%]), corresponding to a rate of incidence of 
`r tbrounding(desc_staff$CT_confirmed_IR)` per 100,000 staff-days [`r tbrounding(desc_staff$CT_symptomatic_IR_lower)`; `r tbrounding(desc_staff$CT_symptomatic_IR_upper)`].

**Interpretation** Despite uncertain case ascertainment in care homes, internal surveillance suggests a greater transmission than in the community. Staffing may impact care homes' response to outbreaks and in turn influence rates of confirmed infections The absence of high-quality case ascertainment hinders local infection prevention initiatives which could be better supported by laboratory data surveillance.

**Funding:** This work was supported by the Economic and Social Research Council [grant reference ES/V003887/1].

# Background

Care home residents are extremely vulnerable to infection due to their advanced age and high prevalence of comorbidity, and this is compounded by frequent exposure to infection, for example through contact with infected staff members. At the peak of the pandemic the number of deaths in UK care home residents was threefold higher than the equivalent period in 2019,[@onsDR2020] but estimates of infection incidence and prevalence over this time period are lacking due to delays in establishing universal testing for care home residents and staff.

For example, at the start of the pandemic, testing for SARS-CoV-2 infection testing using reverse transcriptase polymerase chain reaction (RT-PCR) was only routinely available for care home residents or staff who were admitted to hospital. Testing was also performed in care homes that notified an outbreak to Public Health England, but this was limited to a maximum of five test per care home.  Widespread testing for residents first became available on 11 May 2020 through the whole care home testing programme. [@dhscTesting2020] Consequently national estimates of incidence and prevalence based on test results obtained prior to 11 May will substantially underestimate the burden of infection in care home residents.

An important source of epidemiological data on the burden of disease in care homes are outbreak investigations.[@McMichael2020; @Dora2020] A living systematic review[@Salcher-Konrad2020] of outbreak investigations found incidence proportions ranging between 0% and 72% among residents and between 2% and 64% among staff of long term care facilities. Estimates of this nature cannot be generalised due to the unknown proportion of care homes that have not reported any outbreaks.

Information on disease incidence and prevalence, and how this varies by care home is essential to inform the public health response to future waves of infection, and in particular decisions around the need for widespread universal testing in care home staff and residents.

This study obtained data from `r formatbm(dplyr::n_distinct(residents$home_code))` care homes from the Four Seasons Health Care Group (FSHCG), one of the UK's largest independent provider of long-term and respite residential and nursing care. It estimated incidence of SARS-CoV-2 symptoms, diagnoses and all-cause mortality. In addition, it tested and identified individual and organisational risk factors associated with the rate of confirmed infections.

# Methods

In 2020, FSHCG represented approximately 9% of all beds in England, Scotland and Northern Ireland (`r table_nums("tab_national_descriptives", display = "cite")`). As of 22 June 2020, 90% of FSHCG care homes have participated in the whole care home testing programme.

FSHCG collects administrative data on their private residents through a variety of electronic systems that are primarily used for billing purposes. This includes individual-level data on the dates of care home entry and exit for residents, demographic information, and reports of infection through an incidents management system known as "Datix". Such records do not include residents occupying beds contracted by local authorities as part of block agreements.

`r table_nums("tab_national_descriptives")` [@cqcdata2020; @hscni2020; @isd2020]

```{r table_national_descriptives, fig.cap = tab_national_descriptives, message=F}
mutate_all(desc_homes,  .funs = formatbm) %>% arrange(gsub("FSHC ", "", Scope)) %>% t %>% kable
```

## Aggregate outbreak data (`r format(date_min_tallies, "%d/%m/%Y")`--`r format(date_max_tallies, "%d/%m/%Y")` )

On 24 March 2020, FSHCG introduced a bespoke reporting system to monitor COVID-19 cases in each care home. This new system requires care home managers to report daily tallies of symptomatic and confirmed infections identified among staff and residents to their FSHCG regional manager.

Daily counts reported by care homes to FSHCG were obtained for every care home:

- new symptomatic cases (residents)
- new confirmed cases in home (residents)
- new confirmed cases in hospital (residents)
- deaths related to COVID-19 (Haydn please confirm what these correspond to: suspected/coroner report/tested positive before death?)
- new symptomatic cases (staff)
- new confirmed cases (staff)
- deaths related to COVID-19 (staff)

In addition, the number of beds currently occupied was reported every Sunday. This weekly information was used to approximate daily occupancy using linear interpolation.

These counts include events relating to residents occupying local authority contract beds as well as a minority of residents receiving a different category of support (eg. young people) who were excluded from the study population.

## Individual-level residents data (`r format(date_min, "%d/%m/%Y")`--`r format(date_max, "%d/%m/%Y")` )

Secondary use individual-level data were extracted and pseudonymised by HW. They consisted of two datasets:

- **a residents dataset** containing one record per resident per care home, alongside residents' gender, date of birth, most recent dates of admission and discharge, date of first admission, type of stay (residential/nursing) and care (general, dementia, elderly). This dataset does not include any records for occupants of local authority 'block contract' beds  ($n=$ `r formatbm(sum(desc_homes[["Total contract beds"]], na.rm=T))`).
- **an incidents dataset** containing `r formatbm(desc_before_exclusions$nrow_incidents)` reports filed by care home staff into FSHCG's Datix incidents management system. Mandated report field were: resident forename, surname, care home identifier, incident date/time, and date/time of reporting. In addition, reports can include: date of birth, gender, information on COVID-19 symptoms, tests and test results, resident current location (home/hospital), and death.

HW created a resident index file bridging the two datasets using: the first three letters of the resident's forename; the resident's full surname; and the care home identifier. Clerical review of identifiers generated in this way established that no single identifier was allocated to distinct individuals in the residents dataset. Just `r desc_before_exclusions$missing_id` (`r tbrounding(desc_before_exclusions$missing_id_pct)`%) Datix reports did not contain enough information to generate an identifier and were excluded. Identifiers of a further `r desc_before_exclusions$link_fail` reports (`r desc_before_exclusions$link_fail_pct`%) could not be successfully linked to the residents dataset and were excluded, after clerical review established they could not be matched with a resident. A proportion of these reports may have been filed in error, while others may relate to occupants of local authority beds who are not accounted for in the residents dataset.

## Analysis

The two sources of data were employed for separate purposes.

Incidence proportions and rates were calculated using the aggregate outbreak counts, since they were the trusted source of information and encompass all residents. The number of unique residents $N$ present in the homes at some point between `r format(date_min_tallies, "%e %B %Y")` and `r format(date_max_tallies, "%e %B %Y")`  was estimated by multiplying the number of private residents (from the individual-level dataset) by the ratio of the total resident-days (from the weekly occupancy censuses) to the total private resident-days (from the individual data). This approach assumes that the mean stay of local authority residents is identical to that of private residents (supplementary material 1). As the total exposed-to-risk (susceptible resident-days) was unknown, it was approximated in a multiple decrement life table by making the assumption that previously symptomatic/confirmed residents were discharged at the same rate as susceptible residents, as was suggested by the individual-level records. Admitted cases were accounted for in the daily new case count. Based on this $\hat{E_t}$, the estimated number or previously symptomatic/confirmed residents present in the home on day $t$ was computed as

$$\hat{E_t} = min(n_{t}/n_{t-1}, 1)( \hat{E}_{t-1} + d_{t-1}); \qquad  \qquad  \qquad E_{t_0} = 0$$
where $d_{t}$ denotes the number of new cases for each day $t = \{1, \ldots, T\}$, while $min(n_{t}/n_{t-1}, 1)$ acts as an (underestimated) approximation of the discharge rate. The total exposure time was calculated as $ \sum_{t=1}^T (N_t - \hat{E_t})$ where $N_t$ denotes daily occupancy figures. The incidence proportion of confirmed cases in England was contrasted with incidence measured from a sample survey of English private households which tested nose/throat swabs from 26,995 individuals between 26 April and 13 June 2020.[@cisbulletin2020; @cisdata2020]

As for individual-level data, they were known to be subject to underreporting of infection cases as well as to exclude residents occupying local authority contracted beds. In spite of this, individual-level provided an accurate of the duration of exposure in resident-days as well as right-censoring, since individual records indicate when a resident was discharged or died. By assuming underreporting of cases to affect residents conditionally at random, it was possible to use the data to study factors associated with the rate of infection and fatalities. A Cox proportional hazards model was used to test the effect of both individual-level risk factors (sex, age, care type) and organisation-level risk factors (region, care home size, staff-to-bed ratio, mean bed occupancy proportion). Dates of death were used to estimate the case-fatality rate, as well as compare all-cause mortality based on the presence of symptoms and infection confirmation. A time-dependent Cox proportional hazards model analysed the effect of the same factors on all-cause mortality based on the time-dependent infection classification (susceptible; symptomatic not confirmed; asymptomatic confirmed; symptomatic confirmed).

We also investigated whether the number of confirmed infections among staff was associated with subsequent infections amongst residents, independently of time. For this, a Poisson model regressed the rate of confirmed resident cases against home occupancy levels as well as lagged rates of confirmed staff infections at care home level.

Data were analysed in `R`3·5·0[@RCoreTeam2018] using the `epitool`[@Aragon2020] and `survival`[@therneau2015] libraries. All computer syntax is available on an online repository.[@Dutey2020] Ninety-five percent confidence intervals for proportions and rates were computed from the exact Poisson and binomial limits.

# Findings

## Descriptives

The study population consisted of an estimated `r formatbm(desc_residents$CT_total_residents)` residents across England, Scotland and Northern Ireland. This included `r formatbm(desc_residents$DAT_total_residents)` (`r tbrounding(desc_residents$DAT_total_residents/desc_residents$CT_total_residents*100)`%) private residents, the vast majority (85·4%) of whom were permanent residents. Characteristics of private residents could be described from individual-level records (`r table_nums("tab_resident_descriptives", display = "cite")`); `r tbrounding(mean(residents$home_type=="Nursing")*100)`% were in nursing homes, and 39·2% received dementia care.

`r table_nums("tab_resident_descriptives")`

```{r table_resident_descriptives, message = FALSE, results="asis"}
arsenal::tableby(home_type  ~ gender + ageg  +
   user_type2 + admission_type2 + covid_status + status + region_UK, 
   control = arsenal::tableby.control(test = FALSE, numeric.stats = c("Nmiss", "meansd"), digits = 1L),
   data = residents_overview) %>% summary()
```

### Resident cases

Between `r format(date_min_tallies, "%e %B %Y")` and `r format(date_max_tallies, "%e %B %Y")`, care homes reported `r formatbm(sum(new_cases$new_symptomatic_home, na.rm = T))` cases of residents developing symptoms in homes, alongside `r formatbm(sum(new_cases$new_symptomatic_hospital, na.rm = T))` residents readmitted from hospitals presenting symptoms, contributing to an overall incidence proportion of `r tbrounding(desc_residents$CT_p_symptomatic)`% [`r tbrounding(desc_residents$CT_p_symptomatic_lower)`%; `r tbrounding(desc_residents$CT_p_symptomatic_upper)`%] or an incidence rate of `r tbrounding(desc_residents$CT_IR_symptomatic)` per 100,000 [`r tbrounding(desc_residents$CT_IR_symptomatic_lower)`; `r tbrounding(desc_residents$CT_IR_symptomatic_upper)`] (`r table_nums("supp_epi_CT", display = "cite")`).

`r table_nums("supp_epi_CT")`

|               |  Symptomatic  | Residents  <br> Confirmed  | Symptomatic | Staff <br> Confirmed   |
| ------------- | -------------:| -------------:|-------------:|-------------:|
| Cases      | `r formatbm(desc_residents$CT_total_symptomatic)` |	`r formatbm(desc_residents$CT_total_confirmed)` |	`r formatbm(desc_staff$CT_symptomatic_total)`|	`r formatbm(desc_staff$CT_confirmed_total)` |
|	$N$ exposed  | `r formatbm(desc_residents$CT_total_residents)`|	`r formatbm(desc_residents$CT_total_residents)`|	`r formatbm(desc_staff$total_staff)`|	`r formatbm(desc_staff$total_staff)`|
|	Total exposure (days)| `r formatbm(desc_residents$CT_total_symptomatic_denom)`	|`r formatbm(desc_residents$CT_total_confirmed_denom)`	|`r formatbm(desc_staff$CT_symptomatic_denom)`	|`r formatbm(desc_staff$CT_confirmed_denom)`|
|	Incidence proportion (%)  | `r tbrounding(desc_residents$CT_p_symptomatic)` [`r tbrounding(desc_residents$CT_p_symptomatic_lower)`; `r tbrounding(desc_residents$CT_p_symptomatic_upper)`]| `r tbrounding(desc_residents$CT_p_confirmed)` [`r tbrounding(desc_residents$CT_p_confirmed_lower)`; `r tbrounding(desc_residents$CT_p_confirmed_upper)`] | 	`r tbrounding(desc_staff$CT_symptomatic_p)` [`r tbrounding(desc_staff$CT_symptomatic_p_lower)`; `r tbrounding(desc_staff$CT_symptomatic_p_upper)`]	| `r tbrounding(desc_staff$CT_confirmed_p)` [`r tbrounding(desc_staff$CT_confirmed_p_lower)`; `r tbrounding(desc_staff$CT_confirmed_p_upper)`] |
|	Incidence rate (per 100,000) | `r tbrounding(desc_residents$CT_IR_symptomatic)` [`r tbrounding(desc_residents$CT_IR_symptomatic_lower)`; `r tbrounding(desc_residents$CT_IR_symptomatic_upper)`] |`r tbrounding(desc_residents$CT_IR_confirmed)` [`r tbrounding(desc_residents$CT_IR_confirmed_lower)`; `r tbrounding(desc_residents$CT_IR_confirmed_upper)`] | `r tbrounding(desc_staff$CT_symptomatic_IR)` [`r tbrounding(desc_staff$CT_symptomatic_IR_lower)`; `r tbrounding(desc_staff$CT_symptomatic_IR_upper)`]| `r tbrounding(desc_staff$CT_confirmed_IR)` [`r tbrounding(desc_staff$CT_symptomatic_IR_lower)`; `r tbrounding(desc_staff$CT_symptomatic_IR_upper)`] |

FSHCG residents underwent reverse transcriptase polymerase chain reaction (RT-PCR) viral antigen diagnostics based on nose and throat swab antigen tests during hospitalisations in the first instance, but such tests later became available in care homes from ***XXXXX***. From 11 May 2020, most FSHCG home undertook testing every resident as part of the whole home testing programme.[@dhscTesting2020] Between `r format(date_min_tallies, "%e %B %Y")` and `r format(date_max_tallies, "%e %B %Y")`  `r formatbm(sum(new_cases$new_confirmed_home, na.rm = T))` residents had an infection confirmed in homes while `r formatbm(sum(new_cases$new_confirmed_hospital, na.rm = T))` had an infection confirmed in hospital. The overall confirmed infection incidence proportion was `r tbrounding(desc_residents$CT_p_confirmed)`% [`r tbrounding(desc_residents$CT_p_confirmed_lower)`%; `r tbrounding(desc_residents$CT_p_confirmed_upper)`%, with an incidence rate of `r tbrounding(desc_residents$CT_IR_confirmed)` per 100,000 [`r tbrounding(desc_residents$CT_IR_confirmed_lower)`; `r tbrounding(desc_residents$CT_IR_confirmed_upper)`].

Kaplan-Meier estimates of the temporal trend in symptomatic and confirmed cases (`r fig_nums("fig_KM_CT_residents_symptomatic", display = "cite")`, `r fig_nums("fig_KM_CT_residents_confirmed", display = "cite")`) indicate wide variation across the UK. We note a strong progression of infections in the North West, North East and South East of England, where more than 1 in 5 residents had a confirmed infection. The sharp increase in confirmed infections in the North West between 30 April and 2 May (67 new cases) can largely be attributed to the piloting of whole-home testing in one care home, leading to 42 cases being confirmed.

```{r KM_CT_residents_symptomatic, fig.height = 4, fig.width = 8}
order_key <-km_ct_residents_symp %>% 
  group_by(region_UK) %>% 
  summarise(S_t = min(S_t)) %>% 
  arrange(S_t)
km_ct_residents_symp$region_UK <- factor(
  km_ct_residents_symp$region_UK,
  levels = order_key$region_UK
)
km_ct_residents_symp$region_group <- cut(as.integer(km_ct_residents_symp$region_UK), 2, labels = 1:2)

ggplot(aes(x = date, y = 1-S_t, 
           ymax = 1 - St_max,
           ymin = 1 - St_min,
           color = region_UK), data = km_ct_residents_symp) +
  # facet_grid(region_group ~ .) +
  geom_ribbon(alpha=0.05, linetype = 2,stat = ) +
  geom_path(lwd=1.2) +
  scale_y_continuous("Cum probability symptomatic case",
                     breaks = seq(0,1,.05)) +
  scale_color_discrete("Region") +
  scale_x_date(name = "Date", 
               breaks = unique(get_monday_date(km_ct_residents_symp$date)),
               labels = format.Date(
                 unique(get_monday_date(km_ct_residents_symp$date)),
                 "%d/%m/%y"
               )) +
  theme_bw()  +
  theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1))

```

`r fig_nums("fig_KM_CT_residents_symptomatic")`

```{r KM_CT_residents_infection, fig.height = 4, fig.width = 8}
order_key <-km_ct_residents_confirmed %>% 
  group_by(region_UK) %>% 
  summarise(S_t = min(S_t)) %>% 
  arrange(S_t)
km_ct_residents_confirmed$region_UK <- factor(
  km_ct_residents_confirmed$region_UK,
  levels = order_key$region_UK
)
km_ct_residents_confirmed$region_group <- cut(as.integer(km_ct_residents_confirmed$region_UK), 2, labels = 1:2)

ggplot(aes(x = date, y = 1-S_t, 
           ymax = 1 - St_max,
           ymin = 1 - St_min,
           color = region_UK), data = km_ct_residents_confirmed) +
  # facet_grid(region_group ~ .) +
  geom_ribbon(alpha=0.05, linetype = 2,stat = ) +
  geom_path(lwd=1.2) +
  scale_y_continuous("Cum probability confirmed case",
                     breaks = seq(0,1,.02)) +
  scale_color_discrete("Region") +
  scale_x_date(name = "Date", 
               breaks = unique(get_monday_date(km_ct_residents_confirmed$date)),
               labels = format.Date(
                 unique(get_monday_date(km_ct_residents_confirmed$date)),
                 "%d/%m/%y"
               )) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1))

```

`r fig_nums("fig_KM_CT_confirmed_residents")`

For England only, a total of `r england_results$n_positive` confirmed cases occurred out of a total `r formatbm(england_results$n_positive + england_results$n_negative)` residents. In comparison, the survey of English community households[@cisbulletin2020] found a total of `r cis_data$n_positive` confirmed cases out of `r formatbm(cis_data$n_positive + cis_data$n_negative)` sampled individuals. This implies a relative risk of `r tbrounding(england_results$RR_england)` [`r tbrounding(exp(log(england_results$RR_england) - 1.96 * england_results$RR_SE)) `; `r tbrounding(exp(log(england_results$RR_england) + 1.96 * england_results$RR_SE))`].

Incidence proportions and rates were also estimated for private residents only using the Datix incident reports dataset (supplementary material 1-2). Estimates are lower than corresponding estimates based on outbreak counts. This is thought to be caused by a combination of underascertainment as well as linkage error.

### Staff cases

Out of a total `r formatbm(desc_staff$total_staff)` staff,  `r formatbm(desc_staff$CT_symptomatic_total)`	(`r tbrounding(desc_staff$CT_symptomatic_p)`% [`r tbrounding(desc_staff$CT_symptomatic_p_lower)`%; `r tbrounding(desc_staff$CT_symptomatic_p_upper)`%]) were recorded as having symptoms by FSHCG managers. In total, `r formatbm(desc_staff$CT_confirmed_total)` (	`r tbrounding(desc_staff$CT_confirmed_p)`% [`r tbrounding(desc_staff$CT_confirmed_p_lower)`%; `r tbrounding(desc_staff$CT_confirmed_p_upper)`%]) staff had a confirmed infection by the end of the period. Regional trends (`r fig_nums("fig_KM_staff_symptomatic", display="cite")`, `r fig_nums("fig_KM_staff_confirmed", display="cite")`) indicate a strong progression of infections in the North West, North East and South East of England.

```{r KM_Staff_Symptomatic, fig.height = 4, fig.width = 8}

order_key <-km_staff_symptomatic %>% 
  group_by(region_UK) %>% 
  summarise(S_t = min(S_t)) %>% 
  arrange(S_t)
km_staff_symptomatic$region_UK <- factor(
  km_staff_symptomatic$region_UK,
  levels = order_key$region_UK
)
km_staff_symptomatic$region_group <- cut(as.integer(km_staff_symptomatic$region_UK), 2, labels = 1:2)

ggplot(aes(x = date, y = 1-S_t, 
           ymax = 1 - St_max,
           ymin = 1 - St_min,
           color = region_UK), data = km_staff_symptomatic) +
  # facet_grid(region_group ~ .) +
  geom_ribbon(alpha=0.05, linetype = 2,stat = ) +
  geom_path(lwd=1.2) +
  scale_color_discrete("Region") +  
  scale_y_continuous("Cum probability symptomatic case",
                     breaks = seq(0,1,.05)) +
  scale_x_date(name = "Date", 
               breaks = unique(get_monday_date(km_staff_symptomatic$date)),
               labels = format.Date(
                 unique(get_monday_date(km_staff_symptomatic$date)),
                 "%d/%m/%y"
               )) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1))

```

`r fig_nums("fig_KM_staff_symptomatic")`

```{r KM_Staff_Confirmed, fig.height = 4, fig.width = 8}

order_key <- km_staff_confirmed %>% 
  group_by(region_UK) %>% 
  summarise(S_t = min(S_t)) %>% 
  arrange(S_t)
km_staff_confirmed$region_UK <- factor(
  km_staff_confirmed$region_UK,
  levels = order_key$region_UK
)
km_staff_confirmed$region_group <- cut(as.integer(km_staff_confirmed$region_UK), 2, labels = 1:2)

ggplot(aes(x = date, y = 1-S_t, 
           ymax = 1 - St_max,#(S_t + 1.96 * S_t_SE),
           ymin = 1 - St_min, #(S_t - 1.96 * S_t_SE),
           color = region_UK), data = km_staff_confirmed) +
  # facet_grid(region_group ~ .) +
  geom_ribbon(alpha=0.05, linetype = 2,stat = ) +
  geom_path(lwd=1.2) +
  scale_color_discrete("Region") +
  scale_y_continuous("Cum probability confirmed case",
                     breaks = seq(0,1,.01)) +
  scale_x_date(name = "Date", 
               breaks = unique(get_monday_date(km_staff_confirmed$date)),
               labels = format.Date(
                 unique(get_monday_date(km_staff_confirmed$date)),
                 "%d/%m/%y"
               )) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1))

```

`r fig_nums("fig_KM_staff_confirmed")`

### Outbreak homes

Between `r format(date_min_tallies, "%e %B %Y")` and `r format(date_max_tallies, "%e %B %Y")`, `r desc_residents$CT_zero_case`/`r desc_residents$total_homes` homes (`r tbrounding( desc_residents$CT_zero_case/desc_residents$total_homes*100)`%) care homes had not reported a single confirmed infection, whether acquired in the home or in hospital (`r table_nums("tab_zero_cases", display="cite")`).

`r table_nums("supp_epi_CT_oubreak_only")`

|               |  Symptomatic  | Residents  <br> Confirmed  | 
| ------------- | -------------:| -------------:|
| Cases      | `r formatbm(desc_residents_outbreak_only$CT_total_symptomatic)` |	`r formatbm(desc_residents_outbreak_only$CT_total_confirmed)` |
|	$N$ exposed  | `r formatbm(desc_residents_outbreak_only$CT_total_residents)`|	`r formatbm(desc_residents_outbreak_only$CT_total_residents)`|	
|	Total exposure (days)| `r formatbm(desc_residents_outbreak_only$CT_total_symptomatic_denom)`	|`r formatbm(desc_residents_outbreak_only$CT_total_confirmed_denom)`
|	Incidence proportion (%)  | `r tbrounding(desc_residents_outbreak_only$CT_p_symptomatic)` [`r tbrounding(desc_residents_outbreak_only$CT_p_symptomatic_lower)`; `r tbrounding(desc_residents_outbreak_only$CT_p_symptomatic_upper)`]| `r tbrounding(desc_residents_outbreak_only$CT_p_confirmed)` [`r tbrounding(desc_residents_outbreak_only$CT_p_confirmed_lower)`; `r tbrounding(desc_residents_outbreak_only$CT_p_confirmed_upper)`] | 
|	Incidence rate (per 100,000) | `r tbrounding(desc_residents_outbreak_only$CT_IR_symptomatic)` [`r tbrounding(desc_residents_outbreak_only$CT_IR_symptomatic_lower)`; `r tbrounding(desc_residents_outbreak_only$CT_IR_symptomatic_upper)`] |`r tbrounding(desc_residents_outbreak_only$CT_IR_confirmed)` [`r tbrounding(desc_residents_outbreak_only$CT_IR_confirmed_lower)`; `r tbrounding(desc_residents_outbreak_only$CT_IR_confirmed_upper)`] |

## Factors associated with confirmed infections

Using the individual-level data on private bed residents (`table_nums("tab_resident_descriptives", display = "cite")`), factors affecting the rate of confirmed cases were investigated in a Cox Proportional Hazard model (`r table_nums("tab_CoxPH_confirmed", display = "cite")`). According to this, and other than geographical location, the staff-to-bed ratio was the only factor significantly associated with the rate of confirmed cases. An increase by 10 percentage point in the staff headcount relative to the total bed count was associated with a 13% reduction [9%; 18%] in the rate of confirmed cases.

`r table_nums("tab_CoxPH_confirmed")`

```{r CPH_confirmed,   results="asis"}
cph_conf <- coxph(Surv(time = start_int, time2 = stop_int +1, event = confirmed) ~ gender + ageg+home_type+user_type2+region_UK+ staff_to_bed_ratio+bed_tot_category + scale(pct_occupancy, scale = F), data = residents_overview)
 summary(cph_conf)
# cox.zph(cph_conf)
# survfit(Surv(time = start_int, time2 = stop_int +1, event = confirmed)~factor(gender), data = residents_overview) %>% ggsurvplot(ylim=c(.85,1), conf.int = T, legend = "right")
explanatory = c("gender", "ageg", "home_type", "user_type2", "region_UK" ,  
                "staff_to_bed_ratio", "bed_tot_category", "pct_occupancy")
dependent = "Surv(time = start_int, time2 = stop_int +1, event = confirmed)"
CPH_confirmed <-  finalfit(residents_overview,
                           dependent, 
                           explanatory, column = T,
                           add_dependent_label = F)  
knitr::kable(CPH_confirmed,
             row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))
```

## Factors associated with all-cause mortality

The crude risk of all-cause death for the entire population of study was 19·5% (`r table_nums("tab_resident_descriptives", display = "cite")`). The all-cause confirmed case fatality rate was `r tbrounding(desc_residents$DAT_CFR)`% [`r tbrounding(desc_residents$DAT_CFR_lower)`%; `r tbrounding(desc_residents$DAT_CFR_upper)`%] based on incident reports. This is to be contrasted with a mortality of `r tbrounding(desc_residents$DAT_BFR)`% [`r tbrounding(desc_residents$DAT_BFR_lower)`%; `r tbrounding(desc_residents$DAT_BFR_upper)`%] in residents that did not test or did not test positive (RR = `r tbrounding(desc_residents$DAT_CFR_RR)` [`r tbrounding(exp(log(desc_residents$DAT_CFR_RR) - 1.96 * desc_residents$DAT_CFR_RR_SE)) `; `r tbrounding(exp(log(desc_residents$DAT_CFR_RR) + 1.96 * desc_residents$DAT_CFR_RR_SE))`]).

The time-dependent Cox proportional hazard models in `r table_nums("tab_CoxPH_death", display = "cite")` examine how mortality is affected by COVID-19 status and other factors. Mortality increased with age and men's rate of death was 46% greater than women's (adjusted HR = 1·5 [1·3; 1·6]) all other things equals. After controlling for other risk factors the hazard ratios of being asymptomatic confirmed and symptomatic confirmed cases respectively were 2·6 [1·6; 4·3] and 9·9 [8·3; 11·9], confirming that symptomatic cases were more severe. As for symptomatic residents without a confirmed diagnosis, the adjusted hazard ratio of all-cause mortality was intermediary at 6·2 [5·1; 7·4]. It is important to note these hazard ratio estimates do not give a comprehensive measure of effect: hazards were not proportional across these categories (`r fig_nums("fig_KM_mortality", display="cite")`). Mortality was significantly higher in nursing care homes (adjusted HR = 1·8 [1·6; 2·0]). The 27% increase in the hazard ratio after covariate adjustment reinforces the finding that this effect does not appear to be confounded by the greater proportion of symptomatic cases or common risk factors. This effect is likely attributable to a greater burden of comorbidities in nursing home residents. Beyond, there was no evidence of an association between care home size and all-cause mortality. After adjusting for other variables, the staff-to-bed ratio did not exhibit a statistically significant association with all-cause mortality (adjusted HR = 0·8 [0·7; 1.1]).

`r table_nums("tab_CoxPH_death")`

```{r risk_death, results="asis"}
# cph_death2 <- coxph(Surv(date1, date2, death_status) ~
#                       gender+ ageg +
#                       staff_to_bed_ratio + bed_tot_category +(region_UK)+ 
#                       home_type*covid_status+ cluster(resident_id), data = residents_episodes)
# summary(cph_death2)
# cox.zph(cph_death2)
# survfit(Surv(time = date1, time2 = date2, event = death_status)~ gender, data = residents_episodes)
# 
explanatory = c("gender", "ageg", "user_type2", "home_type",
                "staff_to_bed_ratio", "bed_tot_category", 
                "covid_status", "region_UK", "cluster(resident_id)")
explanatory_interaction = c("gender", "ageg", "user_type2",
                "staff_to_bed_ratio", "bed_tot_category", 
                "covid_status:home_type", "region_UK", "cluster(resident_id)")
dependent = "Surv(time = date1, time2 = date2, event = death_status)"
CPH_death <-  finalfit(residents_episodes, dependent, explanatory, 
                       column = T, add_dependent_label = F)  
knitr::kable(CPH_death, row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))

```

```{r fig_KM_mortality,  fig.height = 4, fig.width = 8}
KM_mortality_by_covid_status <- survfit(
  Surv(time = date1, time2 = date2, event = death_status) ~ covid_status, 
        data = residents_episodes)
ggsurvplot(KM_mortality_by_covid_status, conf.int = TRUE, 
           legend = "right",
           ggtheme = theme_bw(),  
           break.x.by	= 7,
           break.y.by = .1,
           legend.labs = levels(residents_episodes$covid_status),
           xlab  = paste0("Days since ", format(date_min, "%d %B %Y")),
           ylab = "Survival",
           conf.int.alpha = .1)  
```

*Note*: The SARS-COV-2 classification is time-dependent and applies from the date of incident recorded on the relevant Datix incident report.

`r fig_nums("fig_KM_mortality")`

# Interpretation

This study is to our knowledge the first large scale investigation of the incidence of SARS-CoV-2 infections in UK care homes independently of outbreaks, using individual-level resident records from a large provider network across `r as.integer(date_max-date_min)` days. A comparable study was previously conducted on across 623 care homes in Ontario, Canada.[@Stall2020] Findings confirm a very rapid progression of the epidemic across care home residents, with up to `r  tbrounding(desc_residents$CT_p_symptomatic)`% [`r tbrounding(desc_staff$CT_symptomatic_p_lower)`%; `r tbrounding(desc_staff$CT_symptomatic_p_upper)`%] of residents experiencing symptoms at some stage and `r tbrounding(desc_residents$CT_p_confirmed)`% [`r tbrounding(desc_residents$CT_p_confirmed_lower)`%; `r tbrounding(desc_residents$CT_p_confirmed_upper)`%] receiving a positive nasopharyngeal antigen test result between `r format(date_min, "%e %B %Y")` and `r format(date_max_tallies, "%e %B %Y")`.

Regional variation was noticeable, but inconsistencies between incident reports and manager counts were indicative of data quality issues, particularly in relation to incident reports. Rates of testing across time and care homes were not yet known at the time of writing and likely distorts measures of incidence.

One prominent organisational factor was significantly and independently associated with the risks of confirmed infection: the staff-to-bed ratio at baseline was a protective factor. A 10 percentage point increase was associated with a 13% reduction [9%; 18%] in hazards of confirmed infections. Few organisational variables were available to identify potential confounders or mediators of this association. One hypothesis is that staffing ratios may be a sign of better-resourced care homes (for instance protective personal equipment), and of a greater capability to prescribe and observe effective infection control procedures. Furthermore, staff can become a vector in the transmission of infections. It is possible that higher staffing levels mitigate the transmission of infections since each staff member will care for a lower number of residents. Better staffed homes may also be less reliant on bank or agency staff who work in different locations. A more detailed survey of care home characteristics, along with protocols and measures adopted during the epidemic may help elucidate mechanisms through which staffing levels are associated with lower infection rates.

The case-fatality rate (`r tbrounding(desc_residents$DAT_CFR)`%) was higher than previous literature[@Salcher-Konrad2020], which is not exclusively attributable to the long observation period. An outbreak investigation[@Graham2020] on 4 care homes in London, UK measured a case-fatality rate of 17% among 126 residents over a period of 62 days, while Stall et al.[@Stall2020] measured a rate of 28% in over a period of 53 days. These two studies had a much lower overall mortality incidence proportion at 7% and 6% respectively compared to 19%.

Uncertainty remains regarding the internal validity of estimates presented in this report. It was not possible to determine each care home's outbreak status with confidence. Consequently, incidence was calculated across all homes, unlike the study by Stall et al.[@Stall2020] Beyond uncertainty in the exact number of cases and outbreaks, the viral antigen test campaign is a likely subject of under-ascertainment, on account of both the diagnostics' intrinsically low sensitivity, and of the timing of the systematic test campaign initiated on 11 May 2020. Residents may not have provided sample at their peak viral load, thereby reducing the rate of detection. According to a meta-analysis of 7 studies,[@Kucirka2020] the probability of a false-negative result in an infected person on the day of symptom onset (day 0) was 38% [18%; 65%], decreasing to 20% [12%; 30%] on day 3, subsequently rising to 66% [54%; 77%] on day 16. Consequently findings presented in the current study are likely to be underestimates. This includes estimates of the incidence ratio comparing care homes with the community in England.

Evidence from the USA suggests that frequent testing in care homes can reduce the spread of infection.[@Dora2020] Although increasing testing capacity raises the possibility of undertaking frequent, repeat testing in the care home population, this may not be desirable if the risk of infection is low, since the testing procedure (nasopharyngeal swabs) is invasive, and it may cause distress to vulnerable residents, particularly those with dementia. Understanding which care homes are at greatest risk of outbreaks infection could support efforts to target surveillance and infection prevention accordingly.

# Declarations

## Funding

This work was supported by the Economic and Social Research Council [grant reference ES/V003887/1].

## Research ethics

This study was approved by the University College London Research Ethics Committee (project reference 13355/002).

# Supplementary material

## Supplementary material 1: Quality and methodology report

## Supplementary material 2: Incidence of confirmed and symptomatic cases among private residents according to Datix incident reports.

Datix incidents reports recorded that `r formatbm(desc_residents$unique_residents_tested)` private residents were tested at least once. Test results were known for `r formatbm(desc_residents$unique_residents_test_result)` of them (`r tbrounding(desc_residents$unique_residents_test_result/desc_residents$unique_residents_tested*100)`%), including `r formatbm(desc_residents$unique_residents_test_result_positive)` residents with a positive test result. This data should be treated as representative, as tests coming back negative were not commonly recorded on Datix. Incidence proportions and rates reported in `r  table_nums("supp_epi_DAT", display = "cite")` are lower than corresponding estimates based on outbreak counts. This is thought to be caused by a combination of underascertainment as well as linkage error.

`r table_nums("supp_epi_DAT")`

| Residents     |  Symptomatic  |  Confirmed  | All-cause death | 
| ------------- | -------------:| -------------:|-------------:|
| Cases      | `r formatbm(desc_residents$DAT_total_symptomatic)` |	`r formatbm(desc_residents$unique_residents_test_result_positive)` | `r formatbm(desc_residents$DAT_total_deaths)`  |
|	$N$ exposed  | `r formatbm(desc_residents$DAT_total_residents)`|	`r formatbm(desc_residents$DAT_total_residents)` | `r formatbm(desc_residents$DAT_total_residents)` |
|	Total exposure (days)|`r formatbm(desc_residents$DAT_total_symptomatic_denom)`|	`r formatbm(desc_residents$DAT_total_confirmed_denom)` | `r formatbm(desc_residents$total_denom)` |
|	Incidence proportion (%)  | `r tbrounding(desc_residents$DAT_p_symptomatic)` [`r tbrounding(desc_residents$DAT_p_symptomatic_lower)`; `r tbrounding(desc_residents$DAT_p_symptomatic_upper)`] |`r tbrounding(desc_residents$DAT_p_confirmed)` [`r tbrounding(desc_residents$DAT_p_confirmed_lower)`; `r tbrounding(desc_residents$DAT_p_confirmed_upper)`] |`r tbrounding(desc_residents$DAT_p_death)` [`r tbrounding(desc_residents$DAT_p_death_lower)`; `r tbrounding(desc_residents$DAT_p_death_upper)`] |
|	Incidence rate (per 100,000) | `r tbrounding(desc_residents$DAT_IR_symptomatic)` [`r tbrounding(desc_residents$DAT_IR_symptomatic_lower)`; `r tbrounding(desc_residents$DAT_IR_symptomatic_upper)`] | `r tbrounding(desc_residents$DAT_IR_confirmed)`  [`r tbrounding(desc_residents$DAT_IR_confirmed_lower)`; `r tbrounding(desc_residents$DAT_IR_confirmed_upper)`] | `r tbrounding(desc_residents$DAT_IR_death)` [`r tbrounding(desc_residents$DAT_IR_death_lower)`; `r tbrounding(desc_residents$DAT_IR_death_upper)` ]|

## Supplementary material 3: Additional tables

`r table_nums("tab_zero_cases")`

```{r tab_zero_cases, results="asis"}
home_zero_confirmed %>% 
  transmute(`UK region/nation` = region_UK,
            `Total homes`,
            `Homes with no Datix confirmed case` = paste0(zero_datix, " (", 
                                          tbrounding( pct_datix*100), "%)"),
            `Homes with no manager-reported case` = paste0(zero_CT, " (", 
                                          tbrounding( pct_CT*100), "%)")) %>% 
  bind_rows(
    tibble(`UK region/nation` = "Total",
           `Total homes` = sum(home_zero_confirmed$`Total homes`),
           `Homes with no Datix confirmed case` = paste0(
             sum(home_zero_confirmed$zero_datix), " (", 
             tbrounding( sum(home_zero_confirmed$zero_datix) / 
                           sum(home_zero_confirmed$`Total homes`)*100), "%)"),
            `Homes with no manager-reported case` = paste0(
             sum(home_zero_confirmed$zero_CT), " (", 
             tbrounding( sum(home_zero_confirmed$zero_CT) / 
                           sum(home_zero_confirmed$`Total homes`)*100), "%)"))
  ) %>% kable(align = c("l", "r", "r", "r", "r"))

```

# References