### 2020-07-12 manuscript_appendix/manuscript_appendix.Rmd
manuscript v1.0

---
title: 'Incidence of confirmed COVID-19 cases in care home staff and residents: Data
  quality and methodology report'
bibliography: ../manuscript/references.bib
output:
  html_document:
    code_folding: hide
    df_print: kable
    keep_md: yes
    toc: yes
    toc_depth: 1
  word_document:
    toc: yes
    toc_depth: 1
    reference_docx: ../manuscript/template.docx
always_allow_html: yes
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo=FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 5,
  collapse = TRUE,
  comment = "#>"
)
options(OutDec = ".")

library(summarytools)
st_options(
  headings = FALSE,
  footnote = NA
)
source("../.Rprofile")
devtools::load_all("..")
library(dplyr)
library(ggplot2)
library(DiagrammeR)
library(DiagrammeRsvg)
library(rsvg)
load_data()
date_min <- as.Date("2020-03-02")
date_max <- as.Date("2020-06-07") #as.Date("2020-06-15")
date_min_tallies <- as.Date("2020-03-24")
date_max_tallies = as.Date("2020-06-07")

tline <- dplyr::bind_rows(timelines$timeline)
stopifnot(unique(tline$date_end)>=date_max)
tline <- dplyr::filter(tline, between(date, date_min, date_max)) %>% 
  dplyr::filter(resident_id %in% unique(residents$resident_id))

home_inclusion <- unique(tline$home_code)

reference_homes <- reference_homes %>% 
  dplyr::filter(home_code %in% home_inclusion) %>% 
  dplyr::left_join(dplyr::select(reference_geography, postcode, country, region_UK), by = "postcode")

incidents <- incidents %>% 
  filter(dplyr::between(incident_date, date_min, date_max)) %>% 
  filter(home_code %in% home_inclusion)
beds <- beds %>% 
  filter(home_code %in% home_inclusion)

desc_before_exclusions <- list()
desc_before_exclusions$nrow_incidents <- nrow(incidents)
desc_before_exclusions$missing_id <- sum(grepl("^MISSING", incidents$resident_id))
desc_before_exclusions$missing_id_pct <- mean(grepl("^MISSING", incidents$resident_id))*100
desc_before_exclusions$link_fail <- sum(!incidents$resident_id[
  which(!grepl("^MISSING", incidents$resident_id))] %in% unique(residents$resident_id))
desc_before_exclusions$link_fail_pct <- tbrounding(
  sum(!incidents$resident_id[
    which(!grepl("^MISSING", incidents$resident_id))] %in% unique(residents$resident_id)) / 
    nrow(incidents)*100)

# removing erroneous incident reports (no identifier or not linking to residents file) 
incidents <- filter(incidents, !grepl("^MISSING", resident_id)) %>% 
  filter(resident_id %in% unique(residents$resident_id))

new_cases <- filter(new_cases, between(date, date_min, date_max_tallies)) %>% 
    filter(home_code %in% home_inclusion)

```

# Overview

This report provides additional information on the validation of:

- procedures for classifying the SARS-CoV-2 status (symptomatic, tested, confirmed)
- procedures for computing the at-risk resident-days 
- overall numbers of symptomatic and confirmed cases

# Classification of SARS-CoV-2 cases from incident reports

The SARS-CoV-2 status is determined by a combination of fields from the Datix incidents management system:

- `infection_covid_19_type` is a mandatory field one of two possible values:
  - "Confirmed formal clinical diagnosis of COVID-19"   
  - "Symptoms, but no definite clinical diagnosis of COVID-19"
- `infection_confirmed` contains one of 4 possible values (if not missing):
  - "Yes - positive test result: confirmed case"
  - "Test result not yet received"
  - "No - negative test result"
  - "Not tested"   
- `infection_result` is a free-text field containing, for the most part, the words 'positive' or 'negative'

Incident reports are dynamic: they are designed to be updated as more information becomes available. If all fields are not accurately kept updated, some values may become inconsistent. Inconsistencies are observed in incident reports  as illustrated in the tabulations below. Due to this, a reclassification algorithm is used which prioritises any information on a positive test result.

```{r , results="asis"}
ctable(incidents$infection_confirmed, incidents$infection_covid_19_type, useNA = "always", style = "rmarkdown")
```

## Algorithm

The following algorithm is used to classify symptomatic, tested, and test-confirmed cases from the Datix reports of COVID-19 infections.

```{r echo=FALSE}
DiagrammeR::grViz("../man/figure/case_classification_1.gv", width=700, height= 500) %>% 
  export_svg %>% charToRaw %>% rsvg::rsvg_png("case_classification_1.png", )
knitr::include_graphics("case_classification_1.png")
```

```{r echo = FALSE}
grViz("../man/figure/case_classification_2.gv", width=700, height= 400) %>% 
  export_svg %>% charToRaw %>% rsvg::rsvg_png("case_classification_2.png")
knitr::include_graphics("case_classification_2.png")
```

```{r echo = FALSE}
DiagrammeR::grViz("../man/figure/case_classification_3.gv", width=400, height= 400)  %>% 
  export_svg %>% charToRaw %>% rsvg::rsvg_png("case_classification_3.png")
knitr::include_graphics("case_classification_3.png")
```

## Validation

Data are tabulated below to evaluate the internal consistency of the rules set out in the previous section.

The `infection_result` field provides information on test results which broadly agrees yet sometimes conflicts with the `infection_confirmed` classification of incidents. It is potentially more up to date although it is very sparsely populated.

```{r, results="asis"}
ctable(incidents$infection_covid_19_type, incidents$covid_confirmed, useNA = "always", style = "rmarkdown")
```

```{r, results="asis"}
ctable(incidents$infection_covid_19_type, incidents$covid_test_result, useNA = "always", style = "rmarkdown")
```

```{r, results="asis"}
ctable(incidents$infection_confirmed, incidents$covid_test_result, useNA = "always", style = "rmarkdown")
```

# Data quality

## Data linkage

Secondary use individual-level data for private residents receiving general and elderly care were extracted and pseudonymised by HW. They consisted of two datasets:

- **a residents dataset** containing one record per resident per care home, alongside residents' gender, date of birth, most recent dates of admission and discharge, date of first admission, type of stay (residential/nursing) and care (general, dementia, elderly). This dataset does not include any records for local authority 'block contract' beds.
- **an incidents dataset** containing `r formatbm(desc_before_exclusions$nrow_incidents)` reports filed by care home staff. Mandated report field were: resident forename, surname, care home identifier, incident date/time, and date/time of reporting. In addition, reports could record: date of birth, gender, information on Covid-19 symptoms, tests and test results, resident current location (home/hospital), and death. Incidents were classified to indicate whether the resident was symptomatic, and whether an infection was confirmed on the basis of multiple variables (supplementary material 1).

HW created a resident index file bridging the two datasets using: the first three letters of the resident's forename; the resident's full surname; and the care home identifier. Clerical review of identifiers generated in this way established that no single identifier was allocated to distinct individuals in the residents' dataset. Just `r desc_before_exclusions$missing_id` (`r tbrounding(desc_before_exclusions$missing_id_pct)`%) incident records did not contain enough information to generate an identifier and were excluded. Identifiers of a further `r desc_before_exclusions$link_fail` records (`r desc_before_exclusions$link_fail_pct`%) could not be successfully linked to the residents dataset and were excluded, after clerical review established they could not be matched with a resident. A proportion of these records may have been filed in error, while others may relate to occupants of local authority beds who are not accounted for in the residents' dataset.

## Exclusions

A total of `r desc_before_exclusions$nrow_incidents` incident reports were obtained. Resident pseudoidentifiers were absent in `r formatbm(desc_before_exclusions$missing_id)` (`r tbrounding(desc_before_exclusions$missing_id_pct)`%) of Datix incident reports. The pseudoidentifiers of a further `r desc_before_exclusions$link_fail ` reports (`r tbrounding(desc_before_exclusions$link_fail /desc_before_exclusions$nrow_incidents*100)`%) were could not be successfully linked to the residents database. All were excluded, leaving a total of `r nrow(incidents)` reports dated between `r format(date_min, "%A %e %b %Y")` and  `r format(date_max, "%A %e %b %Y")` relating to `r formatbm(dplyr::n_distinct(incidents$resident_id))` unique residents.
These reports belongued to `r formatbm(dplyr::n_distinct(incidents$home_code))` care homes totalling `r formatbm(sum(beds[beds$year=="2020" & beds$home_code %in% unique(incidents$home_code), "beds"]))` beds. For the period of study, the Four Seasons Health Care Group (FSHCG) comprises `r formatbm(dplyr::n_distinct(beds[beds$year=="2020", "home_code"]))` care homes with a total of `r formatbm(sum(beds[beds$year=="2020", "beds"]))` beds.

## Occupancy

```{r occupancy_appx}
approx_occupancy <- tline %>% 
  filter(date %in% unique(occupancy$week_ending)) %>% 
  mutate(week_ending = date) %>% 
  group_by(home_code, week_ending) %>% 
  summarise(approx_occupancy = sum(rday))

occ_home <- merge(
  approx_occupancy,
  #remove Warmley House Nursing which has no occupancy report
  na.omit(select(occupancy, #filter(occupancy, home_code != "360"),
                 home_code, week_ending, occupancy)),
  all = FALSE
)

bed2020 <- dplyr::filter(beds, year == "2020") %>% 
  dplyr::select(-year) %>% 
  mutate(pct_contract = contract_beds/beds,
         contract_10pc = dplyr::case_when(
    contract_beds/beds >= .1 ~ ">=10% contract beds",
    TRUE ~ "<10% contract beds"
  ))

occ_home <- merge(
  occ_home,
  bed2020,
  all.x = TRUE
) %>% 
  dplyr::mutate(contract_10pc = dplyr::case_when(
    contract_beds/beds >= .1 ~ ">=10% contract beds",
    TRUE ~ "<10% contract beds"
  )) %>% 
  dplyr::left_join(select(reference_homes, home_code, country)) %>% 
 dplyr::filter(between(week_ending, date_min_tallies, date_max_tallies))

ratio_occupant_estimates <- (sum(occ_home$occupancy)/sum(occ_home$approx_occupancy))
ratio_occupant_estimates_england <- dplyr::filter(occ_home, country == "England") %>% 
  dplyr::summarise(ratio = sum(occupancy)/sum(approx_occupancy)) %>% 
  .$ratio

outbreak_homes <- dplyr::mutate(new_cases, new_confirmed = tidyr::replace_na(new_confirmed_home, 0) + tidyr::replace_na(new_confirmed_hospital, 0)) %>% 
  dplyr::group_by(home_code) %>% 
  dplyr::summarise(new_confirmed = sum(new_confirmed)) %>% 
  dplyr::filter(new_confirmed != 0) %>% 
  dplyr::distinct(home_code) %>% .$home_code

ratio_occupant_estimates_outbreak_only <- occ_home %>% 
  dplyr::filter(home_code %in% outbreak_homes) %>% 
  dplyr::summarise(ratio = sum(occupancy)/sum(approx_occupancy)) %>% 
  .$ratio

occ_global <- occ_home %>% 
  dplyr::group_by(home_code, contract_10pc, pct_contract) %>% 
  dplyr::summarise(
    approx_occupancy = sum(approx_occupancy),
    occupancy = sum(occupancy)
  ) %>% 
  dplyr::mutate(
    outliers = dplyr::if_else(
      abs(1-approx_occupancy/occupancy) > .10 &
        contract_10pc != ">=10% contract beds",
      home_code,
      "")
  )
```

Care home resident-days are available from two sources:

- reports to FSHCG by every home on their occupancy every Sunday, which include residents occupying contract beds 
- individual-level resident records of first and last admission and discharge dates, to he exclusion of occupants of contract beds. These records are used to compute occupancy on a daily basis.

In total, `r sum(filter(bed2020, home_code %in% home_inclusion)$contract_beds)` out of `r sum(filter(bed2020, home_code %in% home_inclusion)$beds)` beds were contracted by the local authority, with `r nrow(filter(bed2020, pct_contract>.5 & home_code %in% home_inclusion))` homes having more than half of their beds contracted.

Plots below compare both measures for every Sunday (Figure 1) and in total across the period (Figure 2) at the level of care homes, drawing attention to care homes with a large proportion of contract beds. The Pearson's correlation between the weekly measures was $r=$ `r tbrounding(cor(occ_home$occupancy, occ_home$approx_occupancy), 2)` overall, rising to $r=$ `r with(filter(occ_home, contract_beds/beds <.1), tbrounding(cor(occupancy, approx_occupancy), 2))` once excluding care homes with $\geq$ 10% contract beds. These findings confirm that occupancy could be approximated reliably using just residents' records of first and last admission and discharge dates.

This enabled us to estimate exposure (resident-days) excluding local authority contract beds, since no Datix incident reports were returned for occupants of such beds.

In addition, this validation enabled us to approximate the total number of unique residents (including contract beds) during the period of observation for aggregate count data (`r date_min_tallies` -- `r date_max_tallies`). The ratio of occupants recorded by FSHCG to the number approximated by us was `r ratio_occupant_estimates `. The number of unique residents in non-contract beds was `r formatbm(length(unique(tline$resident_id[tline$rday ==1])))` during the study period. By assuming that the resident turnover in local authority beds was the same as the rate in other beds, we infer the number of unique residents inclusive of contract beds to be `r formatbm(round(length(unique((tline$resident_id[tline$rday ==1])))*ratio_occupant_estimates+.5))`.

In England, this ratio was `r ratio_occupant_estimates_england`, and for outbreak homes across the UK, `r ratio_occupant_estimates_outbreak_only`

```{r,  fig.cap = "Weekly total resident-days per care home", message = F}
ggplot(occ_home, aes(x = occupancy, y = approx_occupancy)) +
  geom_point(aes(shape = contract_10pc)) +
  geom_abline(aes(intercept = 0, slope = 1,  col = "Identity", linetype = "Identity")) +
  geom_smooth(aes(col="OLS", linetype = "OLS"), method = "lm", se = T) +
  xlab("Occupancy provided by FSHC") +
  ylab("Occupancy approximated\nfrom resident records") +
  ggtitle(paste0("Figure 1: Weekly total resident-days by care home (L=", dplyr::n_distinct(occ_home$home_code), ")")) +
  scale_colour_manual(name="Line of fit",
    values = c("Identity" = "black", "OLS" = "blue", "Spline" = "red")) +
  scale_linetype_manual(name="Line of fit",
    values = c("Identity" = "solid", "OLS" = "longdash", "Spline" = "dotted")) +
  scale_shape_manual(name="Bed composition",
    values = c(1, 3)) 
```

```{r, fig.cap = "Weekly total resident-days", message = F}

ggplot(occ_global, aes(x = occupancy, y = approx_occupancy)) +
  geom_point(aes(shape = contract_10pc)) +
  # geom_text(aes(label = outliers), 
  #           nudge_x = .1,  nudge_y = .15, vjust = "outward", hjust = "outward") +
   geom_abline(aes(intercept = 0, slope = 1,  col = "Identity", linetype = "Identity")) +
  geom_smooth(aes(col="OLS", linetype = "OLS"), method = "lm", se = T) +
  xlab("Occupancy provided by FSHC") +
  ylab("Occupancy approximated from resident records") +
  ggtitle(paste0("Figure 2: Total resident-days across FSHC during study period \n(", format(date_min, "%d %b %Y"), " - ", format(date_max, "%d %b %Y"),")")) +
  scale_colour_manual(name="Line of fit",
    values = c("Identity" = "black", "OLS" = "blue", "Spline" = "red")) +
  scale_linetype_manual(name="Line of fit",
    values = c("Identity" = "solid", "OLS" = "longdash", "Spline" = "dotted")) +
  scale_shape_manual(name="Bed composition",
    values = c(1, 3)) 
```

We nevertheless identified 11 care homes in which the total resident-days approximated differed by more than 10 percent despite not belonging (marked on Figure 2).

```{r, message=F}
# merge(dplyr::select(reference_homes, home_code, home_name_datix),
#       dplyr::filter(occ_global, outliers != "")) %>%
#   dplyr::transmute(home_code, home_name_datix, approx_occupancy, occupancy,
#                    `% contract beds` = round(pct_contract*100, 1))
```

## Comparison of manager-reported tally counts with Datix-derived counts

```{r case_validation, message=F}
residents_overview <- tline %>% 
  dplyr::filter(between(date, date_min_tallies, date_max_tallies)) %>% 
  dplyr::filter(rday==1) %>%
  dplyr::group_by(resident_id, home_code) %>% 
  dplyr::summarise(resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed),
            symptomatic = max(-susceptible_symptomatic+1, na.rm = T),
            confirmed = max(-susceptible_confirmed+1, na.rm = T),
            start_date = min(date, na.rm = T),
            stop_date = max(date, na.rm = T),
            start_int = as.numeric(min(date, na.rm = T) - date_min),
            stop_int = as.numeric(max(date, na.rm = T)- date_min)
    ) %>% 
  dplyr::left_join(dplyr::select(reference_homes, home_code, home_name_datix)) %>% 
  dplyr::left_join(dplyr::distinct(dplyr::select(incidents, resident_id, dplyr::starts_with("covid_first"))) )

datix_counts <- residents_overview %>% 
  dplyr::group_by(home_code, home_name_datix) %>% 
  dplyr::summarise(
    datix_symptomatic = sum(symptomatic, na.rm = T),
    datix_confirmed = sum(confirmed, na.rm = T) ) %>% 
  dplyr::left_join(bed2020)

tally_counts <- new_cases %>% 
  filter(home_code %in% home_inclusion) %>% 
  filter(between(date, date_min, date_max)) %>% 
  dplyr::group_by(home_code) %>% 
  dplyr::summarise(
    tally_symptomatic = sum(new_symptomatic_home, na.rm = T) + 
      sum(new_symptomatic_hospital, na.rm = T),
    tally_confirmed = sum(new_confirmed_home, na.rm = T) + 
      sum(new_confirmed_hospital, na.rm = T)
  )

ch_data_all <- merge(datix_counts, tally_counts, all = T) %>% 
  dplyr::mutate(
    home_code_lbl = if_else(tally_symptomatic>20, home_code, ""),
    home_code_lbl2 = if_else(tally_confirmed>15 | datix_confirmed>15, home_code, ""))
```

During the period, manager counted `r sum(ch_data_all$tally_confirmed)` confirmed infection while Datix recorded `r sum(ch_data_all$datix_confirmed)` such infections. The `r round(sum(ch_data_all$tally_confirmed)/sum(ch_data_all$datix_confirmed)*100-100 )`% excess could be due to a range of explanations:

1. contract beds occupants, who may represent at least `r tbrounding((1-1/ratio_occupant_estimates)*100)`% of residents (assuming that their turnover is identical to that of occupants of other beds, in reality turnover is likely to be higher, and so is this proportion), may have a higher infection rate than other residents 
2. underreporting on Datix
3. linkage error
4. double-counting of cases in manager counts

It is not possible to rely on explanations (1-3) exclusively: the number of confirmed infections on Datix was greater than manager counts in `r sum(ch_data_all$datix_confirmed > ch_data_all$tally_confirmed)` care homes. This suggest double counting (4) is a possibility.

```{r comparison_symptomatic, message = F}
ggplot(data = ch_data_all, 
       aes(x = tally_symptomatic, y = datix_symptomatic)) + 
  geom_point(aes(shape = contract_10pc)) +
  # geom_text(aes(label = home_code_lbl),
  #           nudge_x = 2,  nudge_y = 1) + 
  geom_abline(aes(intercept = 0, slope = 1,  col = "Identity", linetype = "Identity")) +
  geom_smooth(aes(col="OLS", linetype = "OLS"), method = "lm", se = T) +
  xlab("Manager counts") + ylab("Datix-derived counts") + labs(title="Figure 3: Comparison of Datix and manager counts of symptomatic cases") +
  scale_colour_manual(name="Line of fit",
    values = c("Identity" = "black", "OLS" = "blue", "Spline" = "red")) +
  scale_linetype_manual(name="Line of fit",
    values = c("Identity" = "solid", "OLS" = "longdash", "Spline" = "dotted")) +
  scale_shape_manual(name="Bed composition",
    values = c(1, 3)) 
```

```{r comparison_confirmed, message=F}

ggplot(data = ch_data_all, aes(x = tally_confirmed, y = datix_confirmed)) + 
  geom_point(aes(shape = contract_10pc)) +
  # geom_text(aes(label = home_code_lbl2), 
  #           nudge_x = 1,  nudge_y = 1.2) + 
  geom_abline(aes(intercept = 0, slope = 1,  col = "Identity", linetype = "Identity")) +
  geom_smooth(aes(col="OLS", linetype = "OLS"), method = "lm", se = T) +
  xlab("Manager counts") + ylab("Datix-derived counts") + labs(title="Figure 4: Comparison of Datix and manager counts of confirmed cases") +
  scale_colour_manual(name="Line of fit",
    values = c("Identity" = "black", "OLS" = "blue", "Spline" = "red")) +
  scale_linetype_manual(name="Line of fit",
    values = c("Identity" = "solid", "OLS" = "longdash", "Spline" = "dotted")) +
  scale_shape_manual(name="Bed composition",
    values = c(1, 3))

```

# Estimating the cumulative incidence of cases and resident-days at risk

Manager-reported outbreak surveillance counts were used together with occupancy estimated from weekly occupied bed censuses in order to estimate the incidence of cases across time using a Kaplan-Meier life table product limit estimator. As the total exposed-to-risk (susceptible resident-days) was unknown, it was approximated in a multiple decrements life table by making the assumption that previously symptomatic/confirmed residents were discharged at the same rate as susceptible residents, as was suggested by the individual-level records. Admitted cases were accounted for in the daily new case count. Based on this $\hat{I_t}$, the estimated number or previously symptomatic/confirmed residents present in the home on day $t$ was computed as

$$\hat{I_t} = min(n_{t}/n_{t-1}, 1)( \hat{E}_{t-1} + d_{t-1}); \qquad  \qquad  \qquad I_{t_0} = 0$$
where $d_{t}$ denotes the number of new cases for each day $t = \{0, \ldots, T\}$, while $min(n_{t}/n_{t-1}, 1)$ acts as an (underestimated) approximation of the discharge rate. The total exposure time was calculated as $\sum_{t=0}^{T} (N_t-\hat{I}_t)$ where $N_t$ denotes daily occupancy figures.

# Total homes, beds, contract beds, and residents within FSHCG and nationwide

Data sources: [@cqcdata2020; @hscni2020; @isd2020]

# References